\chapter[Introduction]{Introduction}\label{cap:intro}

\begin{flushright}
	\textit{''A measurement result is complete only when accompanied \\ 
	by a quantitative statement of its uncertainty. The uncertainty \\
	is required in order to decide if the result is adequate for \\
	its intended purpose and to ascertain if it is consistent with \\
	other similar results'' \\}
\end{flushright}


The rapid growth of high-performance computing and the advances in numerical techniques in the last two decades have provided an unprecedented opportunity to explore complex physical phenomena using large-scale spatio-temporal modeling and simulation.  At the same time, scientific community is leaving behind the traditional deterministic approach, which offers  point predictions with no associated uncertainty \cite{Johnstone2015};  to include Uncertainty Quantification (\textit{UQ}) as a common practice in their researches. 

Large-scale spatio-temporal simulations with quantified uncertainty enable scientists to make precise statements about the degree of confidence they have in their simulation-based predictions. These approaches find practical applicability in models for predicting the behavior of weather, hurricane forecasts \cite{Tobergte2013}, subsurface hydrology \cite{Baroni2014a}, geology \cite{Guerra2016}, nuclear reactor design, financial portfolios \cite{Chen2008}, and biological phenomena, just to name a few. They also allow to study physical phenomena that are impossible to assess experimentally, for example: simulate nuclear accidents, or the conditions that some spatial vehicle will find at landing in Mars, and so on. The success of these techniques has made them increasingly important tools for high impact predictions and decision making.

\textit{UQ}  includes different aspects that warranty the predictive fidelity of a numerical simulation, such as the uncertainty in the experimental data, which is used for defining the parameter values of a model; the propagation of uncertain  parameters through the model; and the choice of the model itself. \textit{UQ} is a complex process that covers the following main tasks: (i) uncertainty characterization \cite{Crespo2014}, also called model calibration \cite{Farrell2015a} or statistical inverse problem \cite{Estacio-Hiroms2012};  (ii) sensitivity analysis; (iii) forward problem or uncertainty propagation; and (iv) model selection.

This thesis is focused on  \textit{forward propagation}, whose objective is to quantify the uncertainties in model output(s) propagated from uncertain inputs. The targets of \textit{forward propagation} analysis can be: (i) evaluate low-order moments ( i.e. mean and variance) of the outputs, (ii) evaluate the reliability of the outputs, and/or (iii) assess the complete probability distribution (\textit{PDF}) of the outputs.

When dealing with large-scale spatio-temporal models, a huge among of data is generated as a result of the simulation process. Indeed, on each spatio-temporal location $(s_{i},t_{j}) \in \mathcal{S} \times \mathcal{T}\subseteq\mathbb{R}^{3}\times\mathbb{R}$ , usually more than $10^4$ simulations are performed. Then, the size of the output dataset is in the order of $N_{s}\times N_{t}\times N_{sim}$, where: $N_{s}$ is the number of spatial locations, $N_{t}$ is the number of time steps, and $N_{sim}$ is the number of simulations.  An example of the volume of data generated by these simulations is given in the first case study of the Chapter \ref{cap:use_cases}, where the output dataset is about 2.4 TB. This turn \textit{forward propagation} in a data intensive problem.

Another important aspect, which is often not taken into account, is that the uncertainty need to be quantified in some way that can be used after, to answer questions that arise in the \textit{UQ} context. In that sense, assess the complete \textit {PDF} could be the best way to quantify uncertainty, because if you can find the \textit {PDF} that best fit the dataset with reasonably accurately, you can get all the statistical properties under one roof. At the same time, we can substitute the original data by the \textit {PDFs}, which represents a huge reduction in the volume of data to manipulate.

Contradictorily, statistical moments (e.g. mean and standard deviation) are possibly the most used ways to quantify the uncertainty, despite the fact that they doesn't have information about the manner in which the data are distributed \cite{Lampasi2006}. This is because of the difficulty to find the \textit{PDF} that best fit a dataset \cite{Karian2011},  even more, when dealing with large-scale spatio-temporal models where the \textit{PDF} needs to be derived on each spatio-temporal location, and therefore the  \textit{forward propagation} problem becomes time consuming and computationally intensive too.

However, the use of low order moments alone prevents us from making accurate analysis with respect to the uncertainty. They are not enough neither for the characterization nor for the quantification of the uncertainty, and questions such as:
\begin{itemize}
\item What is the uncertainty in the spatio-temporal region $\mathcal{S}_{i} \times \mathcal{T}_{j}$ associated to the \textit{QoI} 
$q_{k}$  and a computational model $\mathcal{M}_{m}$?
\item How to compare different spatio-temporal regions $\mathcal{S}_{i} \times \mathcal{T}_{j}$ with respect to the uncertainty? 
\item What is the less uncertain model from the set of models $\mathcal{M}={\mathcal{M}_{1},\mathcal{M}_{2},
....\mathcal{M}_{m}}$, to predict the value of a QoI $q_{k}$, over a spatio-temporal region $\mathcal{S}_{i} \times \mathcal{T}_{j}$?
\end{itemize}
can be poorly answered. So, we emphasize that only the characterization of the uncertainty by using the \textit{PDF} allows aware decisions.

A first effort to try to estimate the \textit{PDFs} on large-scale spatio-temporal simulations was done by \cite{Liu2018} {Ji et. al.} in \textbf{\textit{Parallel Computation of PDFs on Big Spatial Data Using Spark}}. They propose a new solution to efficiently compute the \textit{PDFs} in parallel using Spark, through three methods: data grouping, machine learning prediction and sampling. The main drawback of the proposed approach is that you should try many different distributions, to find the PDF that best fits the dataset at each specific spatio-temporal location. Another drawback is that, as we mentioned above, the uncertainty needs to be quantified in the way that facilitates its further use; and the heterogeneity of the functions used in the approach doesn't facilitate it. 

To face these challenges, in this thesis we propose a general framework to quantify the uncertainty in large-scale spatio-temporal models. It uses a data-driven approach and combines the generalized lambda distribution (\textit{GLD}), clusters algorithms and information entropy, for helping researchers to answer the above questions and many others that arise in \textit{UQ} context. Our proposal provides a generally applicable and easy-to-use tool that supports the representation and analysis of uncertainty, as was suggested in the "\textit{Workshop on Quantification, Communication, and Interpretation of Uncertainty in Simulation and Data Science}" \cite{Tobergte2013}. 

\section{Research Objectives}

\begin{tcolorbox}
The main objective of this thesis is a new method to quantify the uncertainty in large-scale spatio-temporal models based on the Generalized Lambda Distribution (GLD).
\end{tcolorbox}

To achieve that goal the following research questions need to be answered:

\textbf{RQ1.} how to group the output of the UQ process based on the similarity of the uncertainty?

\textbf{RQ2.} what is the uncertainty in some spatio-temporal locations not previously analyzed?

\textbf{RQ3.} what is the uncertainty at a specific spatio-temporal region?

\textbf{RQ4.} how to compare two regions as a function of their uncertainty?

\textbf{RQ5.} what is the least uncertain from a set of models?

\textbf{RQ1} is answered in Chapter \ref{cap:gld_clustering}, while \textbf{RQ2}, \textbf{RQ3}, \textbf{RQ4} and \textbf{RQ5} are answered in Chapter \ref{cap:our_approach}. In Chapter \ref{cap:use_cases} all the questions are answered again on the context of real use cases.

%\section{Highlights of the Dissertation}

\section{Organization of the Dissertation}

The structure of the remainder of this thesis is outlined for reference.

\textbf{Chapter 2} \textbf{[Uncertainty Quantification Background]}. In this Chapter, we review the UQ specialized literature, comment some interesting results and highlight the remainder challenges we are interested in to solving in the present thesis.

\textbf{Chapter 3} \textbf{[The Generalized Lambda Distribution]}. In this Chapter, we review the principal features of the \textit{GLD} that motivated us to select it as a distribution to characterize the uncertainty in our proposed approach. We highlight some interesting aspects of the shapes of the \textit{GLD}, its flexibility to fit a variate well known distributions and datasets in many different fields. Finally we summarize all the characteristics of the \textit{GLD} that make it the perfect candidate to be used in UQ. 

\textbf{Chapter 4} \textbf{[Clustering Uncertain Data Based on GLD Similarity]}. This chapter explore the how the \textit{GLD} can be used to cluster uncertain data. An algorithm was proposed an tested in two synthetic datasets with excellent results. 

\textbf{Chapter 5} \textbf{[Our Approach]}. Based on the challenges highlighted in Chapter \ref{cap:background}, the characteristics of the \textit{GLD} that make it a good candidate to be used in UQ as summarized in Chapter \ref{cap:gld}, and the results of Chapter \ref{cap:gld_clustering}, in this Chapter we present our approach. All the algorithm of the approach are discuss. 

\textbf{Chapter 6} \textbf{[Use Cases]}. In order to illustrate the use of the proposed framework, two case studies are discussed. The main results obtained are: (i) the \textit{GLD} is a good fits for more than the 80\% of the datasets, (ii) the use of the \textit{GLD} allows to include clustering algorithms to group the spatio-temporal locations with similar uncertainty, (iii) the centroids  of the clusters can be used as a faithful representation of the rest of the spatio-temporal locations, which significantly reduces the data corresponding to the simulation outputs, (iv) with the use of these centroids we can characterize the uncertainty in any spatio-temporal region as a mixture of \textit{GLDs}.

\textbf{Chapter 7} \textbf{[Conclusions and Future Works]}. Finally, in this chapter the main contributions of the thesis are highlighted, we revisit the research questions formulated in the Introduction and remark the future directions of this work.
